---
title: "Wasserportal"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Wasserportal}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Define Helper Functions

```{r define_helpers}
library(kwb.geosalz)

add_label <- function(df, columns) {
  df %>%
    dplyr::mutate(label = wasserportal::columns_to_labels(
      data = df,
      columns = columns,
      fmt = "<p>%s: %s</p>",
      sep = ""
    ))
}

add_interval_label <- function(df) {
  add_label(df, c("Nummer", "start", "end", "n", "interval"))
}

add_waterbody_label <- function(df) {
  add_label(df, c("Nummer", "Gewaesser", "Auspraegung", "Betreiber"))
}

get_data_stats <- function(df, group_col = "Messstellennummer") {
  df %>%
    dplyr::group_by(.data[[group_col]]) %>%
    dplyr::summarise(
      start = min(as.Date(.data$Datum)),
      end = max(as.Date(.data$Datum)),
      period = diff(c(.data$start, .data$end)),
      n = dplyr::n(),
      interval = round(.data$period / .data$n)
    )
}

difftime_to_numeric <- function(df) {
  df %>%  
    dplyr::rename(
      period_days = "period", 
      interval_days = "interval"
    ) %>% 
    dplyr::mutate(
      period_days = as.numeric(.data$period_days),
      interval_days = as.numeric(.data$interval_days)
    )
}

remove_na_and_geometry <- function(df) {
  df %>%
    kwb.utils::removeColumns("label") %>%
    kwb.utils::removeEmptyColumns(dbg = FALSE) %>%
    remove_geometry()
}

remove_geometry <- function(df) {
  sf::st_geometry(df) <- NULL
  df
}

write_csv <- function(x, file) {
  kwb.utils::catAndRun(
    sprintf("Writing %s to %s", deparse(substitute(x)), file),
    readr::write_csv2(x, file)
  )
}

# Function to create a map
create_map <- function(
    basemap, data, labels, legend_label, legend_color, circle_color
)
{
  basemap %>%
    leaflet::addCircles(
      data = data,
      color = circle_color,
      opacity = 0.4,
      label = lapply(labels, htmltools::HTML)
    ) %>%
    leaflet::addLegend(
      position = "topright",
      colors = c("red", legend_color),
      labels = c("SVM Modellgrenze", legend_label),
      title = "Legende"
    )
}
```

## Initialise objects

```{r setup}
swl_data_stats_export <- NULL
gwl_data_stats_export <- NULL
gwq_chloride_data_stats_export <- NULL
```

```{r}
# Load SVM boundaries and transform coordinate reference system (CRS)
svm <- "extdata/gis/shapefiles/svm_south.shp" %>%
  system.file(package = "kwb.geosalz") %>%
  sf::st_read() %>%
  sf::st_transform(crs = 4326)

# Print information on the CRS
sf::st_crs(svm)

# Function to filter for study area
filter_for_study_area <- function(x) {
  sf::st_filter(x, y = svm$geometry[svm$name == "Friedrichshagen"])
}

# Get information on "Wasserportal"-stations 
stations <- wasserportal::get_stations()
```

## Download and Process Surface Water Level Data

```{r surfacewater_level}
try({
  
  swl_master <- wasserportal::get_wasserportal_masters_data(
    master_urls = stations$overview_list$surface_water.water_level %>%
      dplyr::filter(.data$Betreiber == "Land Berlin") %>%
      dplyr::pull(.data$stammdaten_link)
  ) 
  
  swl_master_sf <- swl_master %>%
    kwb.geosalz::convert_to_sf() %>%
    filter_for_study_area() %>%
    add_waterbody_label()
  
  swl_master <- swl_master %>% 
    dplyr::filter(.data$Nummer %in% swl_master_sf$Nummer) 
  
  write_csv(swl_master, "swl_master.csv")
  
  swl_data_list <- lapply(
    X = stats::setNames(nm = swl_master$Nummer),
    FUN = wasserportal::read_wasserportal,
    from_date = "1900-01-01",
    variables = "ows", 
    type = "daily",
    stations_crosstable = stations$crosstable
  )
  
  column_level_zero <- "Pegelnullpunkt_m_NHN"
  key_columns <- c("Nummer", column_level_zero)

  swl_data <- stats::setNames(nm = names(swl_data_list)) %>%
    lapply(function(name) swl_data_list[[name]][[1]]) %>% 
    dplyr::bind_rows(.id = "Messstellennummer") %>% 
    dplyr::mutate(Datum = as.Date(.data$Datum, format = "%d.%m.%Y")) %>% 
    dplyr::filter(.data$Tagesmittelwert != -777) %>% 
    dplyr::rename(
      Tagesmittelwert_cm_ueber_Pegelnullpunkt = "Tagesmittelwert"
    ) %>% 
    dplyr::left_join(
      swl_master[, ..key_columns],
      by = c(Messstellennummer = "Nummer")
    ) %>%  
    dplyr::mutate(
      Tagesmittelwert_Pegelstand_mNN = as.numeric(.data[[column_level_zero]]) +
        .data$Tagesmittelwert_cm_ueber_Pegelnullpunkt / 100
    ) %>% 
    kwb.utils::removeColumns(column_level_zero)
  
  write_csv(swl_data, "swl_data.csv")
  
  swl_data_stats <- get_data_stats(swl_data)
  
  swl_data_stats_export <- swl_master_sf %>% 
    remove_na_and_geometry() %>%
    dplyr::left_join(
      difftime_to_numeric(swl_data_stats),
      by = c("Nummer" = "Messstellennummer")
    )
  
  write_csv(swl_data_stats_export, "swl_data_stats.csv")
})
```

## Download and Process Groundwater Level Data

```{r groundwater_level}
try({

  gwl_master <- jsonlite::fromJSON(
    "https://kwb-r.github.io/wasserportal/stations_gwl_master.json"
  )
  
  gwl_master_sf <- gwl_master %>%
    kwb.geosalz::convert_to_sf() %>%
    filter_for_study_area()
  
  gwl_master <- gwl_master %>% 
    dplyr::filter(.data$Nummer %in% gwl_master_sf$Nummer)
  
  write_csv(gwl_master, "gwl_master.csv")
  
  gwl_data <- lapply(
    X = gwl_master$Nummer,
    FUN = wasserportal::read_wasserportal_raw_gw, 
    stype = "gwl"
  ) %>% 
    data.table::rbindlist()
  
  write_csv(gwl_data, "gwl_data.csv")
  
  gwl_data_stats <- get_data_stats(gwl_data) 
  
  gwl_data_stats_export <- gwl_master_sf %>% 
    remove_na_and_geometry() %>%
    dplyr::left_join(
      difftime_to_numeric(gwl_data_stats),
      by = c("Nummer" = "Messstellennummer")
    )
  
  write_csv(gwl_data_stats_export, "gwl_data_stats.csv")
  
  gwl_master_stats <- gwl_master %>%
    dplyr::left_join(gwl_data_stats, by = c(Nummer = "Messstellennummer")) %>%
    add_interval_label()
  
})
```

## Download and Process Groundwater Quality Data

```{r groundwater_quality}
try({
  
  gwq_master <- jsonlite::fromJSON(
    "https://kwb-r.github.io/wasserportal/stations_gwq_master.json"
  )

  gwq_master_sf <- gwq_master %>%
    kwb.geosalz::convert_to_sf() %>%
    filter_for_study_area()
  
  gwq_master <- gwq_master %>% 
    dplyr::filter(.data$Nummer %in% gwq_master_sf$Nummer)
  
  write_csv(gwq_master, "gwq_master.csv")
  
  gw_master <- gwl_master %>%  
    dplyr::full_join(gwq_master) %>% 
    dplyr::mutate(Nummer = as.integer(.data$Nummer)) %>% 
    dplyr::arrange(.data$Nummer)
  
  write_csv(gw_master, "gw_master.csv")
  
  gwq_data <- jsonlite::fromJSON(
    "https://kwb-r.github.io/wasserportal/stations_gwq_data.json"
  ) %>%
    dplyr::filter(.data$Messstellennummer %in% gwq_master$Nummer)
  
  write_csv(gwq_data, "gwq_data.csv")
  
  gwq_chloride_data <- gwq_data %>%
    dplyr::filter(.data$Parameter == "Chlorid")
  
  write_csv(gwq_chloride_data, "gwq_chloride_data.csv")
  
  gwq_chloride_master <- gwq_master %>% 
    dplyr::filter(.data$Nummer %in% gwq_chloride_data$Messstellennummer)
  
  write_csv(gwq_chloride_master, "gwq_chloride_master.csv")
  
  gwq_chloride_data_stats <- gwq_chloride_data %>%
    get_data_stats()
  
  gwq_chloride_data_stats_export <- gwq_chloride_master %>%
    kwb.geosalz::convert_to_sf() %>% 
    remove_na_and_geometry() %>%
    dplyr::mutate(Nummer = as.integer(.data$Nummer)) %>% 
    dplyr::left_join(
      difftime_to_numeric(gwq_chloride_data_stats),
      by = c("Nummer" = "Messstellennummer")
    )
  
  write_csv(gwq_chloride_data_stats_export, "gwq_chloride_data_stats.csv")
  
  gwq_chloride_master_stats <- gwq_chloride_master %>%
    dplyr::mutate(Nummer = as.integer(.data$Nummer)) %>%
    dplyr::left_join(
      gwq_chloride_data_stats, 
      by = c("Nummer" = "Messstellennummer")
    ) %>%
    add_interval_label()
  
  write_csv(gwq_chloride_master_stats, "gwq_chloride_master_stats.csv")
})
```

## Create Maps

```{r print_maps}
swl_map <- NULL
gwl_map <- NULL
gwq_map <- NULL

try({
  
  basemap <- svm %>%
    leaflet::leaflet() %>%
    leaflet::addTiles() %>%
    leaflet::addProviderTiles(leaflet::providers$CartoDB.Positron) %>%
    leaflet::addPolygons(color = "red", fill = FALSE)
  
  swl_map <- create_map(
    basemap,
    data = swl_master_sf, 
    labels = swl_master_sf$label,
    legend_label = "OW-Stand (Wasserportal)",
    legend_color = "darkblue",
    circle_color = "blue"
  )
  
  gwl_map <- create_map(
    basemap,
    data = kwb.geosalz::convert_to_sf(gwl_master_stats),
    labels = gwl_master_stats$label,
    legend_label = "GW-Stand (Wasserportal)",
    legend_color = "blue",
    circle_color = "blue"
  )
  
  gwq_map <- create_map(
    basemap,
    data = kwb.geosalz::convert_to_sf(gwq_chloride_master_stats),
    labels = gwq_chloride_master_stats$label,
    legend_label = "GW-G\u00FCte (Wasserportal)",
    legend_color = "orange",
    circle_color = "orange"
  )
  
  print(swl_map)
  print(gwl_map)
  print(gwq_map)
})

```

## Save Maps 

Save maps for full page view:

```{r save_swl_map, eval = !is.null(swl_map)}
htmlwidgets::saveWidget(
  swl_map, 
  "./map_swl.html", 
  title = "SW level"
)
```

```{r save_gwl_map, eval = !is.null(gwl_map)}
htmlwidgets::saveWidget(
  gwl_map, 
  "./map_gwl.html", 
  title = "GW level"
)
```

```{r save_gwq_map, eval = !is.null(gwq_map)}
htmlwidgets::saveWidget(
  gwq_map, 
  "./map_gwq.html", 
  title = "GW quality (Chloride)"
)
```

**Links to full-page view maps:**

* [https://kwb-r.github.io/kwb.geosalz/map_swl.html](../map_swl.html)

* [https://kwb-r.github.io/kwb.geosalz/map_gwl.html](../map_gwl.html)

* [https://kwb-r.github.io/kwb.geosalz/map_gwq.html](../map_gwq.html)


## Datasets


### Surface Water

* [https://kwb-r.github.io/kwb.geosalz/swl_master.csv](../swl_master.csv)

* [https://kwb-r.github.io/kwb.geosalz/swl_data.csv](../swl_data.csv)

* [https://kwb-r.github.io/kwb.geosalz/swl_data_stats.csv](../swl_data_stats.csv)

### Groundwater

#### Master Data 

for [Groundwater Levels](#groundwater-levels) and [Groundwater Levels](#groundwater-quality)

* [https://kwb-r.github.io/kwb.geosalz/gw_master.csv](../gw_master.csv)

#### Groundwater Levels

* [https://kwb-r.github.io/kwb.geosalz/gwl_master.csv](../gwl_master.csv)

* [https://kwb-r.github.io/kwb.geosalz/gwl_data.csv](../gwl_data.csv)

* [https://kwb-r.github.io/kwb.geosalz/gwl_data_stats.csv](../gwl_data_stats.csv)

#### Groundwater Quality 

* [https://kwb-r.github.io/kwb.geosalz/gwq_master.csv](../gwq_master.csv)

* [https://kwb-r.github.io/kwb.geosalz/gwq_data.csv](../gwq_data.csv)


* [https://kwb-r.github.io/kwb.geosalz/gwq_chloride_master.csv](../gwq_chloride_master.csv)

* [https://kwb-r.github.io/kwb.geosalz/gwq_chloride_master_stats.csv](../gwq_chloride_master_stats.csv)

* [https://kwb-r.github.io/kwb.geosalz/gwq_chloride_data.csv](../gwq_chloride_data.csv)


### Tables

#### SW Levels 

```{r tbl_swl, eval = !is.null(swl_data_stats_export)}
DT::datatable(swl_data_stats_export)
```

#### GW Levels 

stats for `groundwater levels`

```{r tbl_gwl, eval = !is.null(gwl_data_stats_export)}
DT::datatable(gwl_data_stats_export)
```

#### GW Quality 

stats for `Chloride`

```{r tbl_gwq, eval = !is.null(gwq_chloride_data_stats_export)}
DT::datatable(gwq_chloride_data_stats_export)
```
